---
title: "Retail"
author: "Sofola Folakunmi Abidemi"
date: "2024-06-27"
output: html_document
---

## INTRODUCTION TO THE PROJECT
This project aim to analyze retail transctional data to gain insight into customer behavior, product perform and sales trends.The gold is to 
help walmart(assuming the company name)make informed decision to grow and make more sales. 

##ASK PHASE
This ask pahse involves the business promblem the project will slove, and there incule;
1. What are the cusmtomer segment and their behaviour? 
2. which product and the most popular among different customer segment?
3. what are the sales trends over time?
4. how do shipping and payment method affect customer over time?

## PREPARE PHASE
Install and load the packages
```{r}
library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
```

Reading .csv into R
```{r}
retail <- read_csv ("C:/Users/USER/Desktop/BACKUP/Desktop/DATA ANALYSIS CLASS/CIRCVEE/new_retail_data.csv")
```

```{r}
head(retail)
```
```{r}
spec(retail)
```

```{r}
dim(retail)
```

```{r}
summary(retail)
```

## PROCESS PHASE 
```{r}
retail_clean <- na.omit (retail)
```

```{r}
summary(retail_clean)
```
### to remove null values
```{r}
retailc <- retail %>% drop_na()
```

```{r}
summary(retailc)
```
### This is to replace null values with the number of mean in a coloumn, in this example the coloum is the AMOUNT. 
```{r}
retail$Amount[is.na(retail$Amount)]<- mean (retail$Amount,na.rm=TRUE)
```

```{r}
summary(retail)
```
```{r}
retail$Ratings[is.na(retail$Ratings)]<- mean (retail$Ratings, na.rm=TRUE)
```

```{r}
summary(retail)
```
### This is to check the mode, mode_impute is naming the fuction, 
```{r}
mode_impute <- function(x) {
  ux <- unique(x)
ux[which.max(tabulate(match(x, ux)))]
}

retail$Zipcode[is.na(retail$Zipcode)] <- mode_impute(retail$Zipcode)
```

```{r}
summary(retail)
```

```{r}
mode_impute <- function(x) {
  ux <- unique(x)
ux[which.max(tabulate(match(x, ux)))]
}

retail$Year[is.na(retail$Year)] <- mode_impute(retail$Year)
```

```{r}
summary(retail)
```

```{r}
head(retail)
```

```{r}
retailc$Amount[is.na(retailc$Amount)]<- mean (retailc$Amount,na.rm=TRUE)
```

```{r}
summary(retailc)
```

```{r}
mode_impute <- function(x) {
  ux <- unique(x)
ux[which.max(tabulate(match(x, ux)))]
}

retailc$year[is.na(retailc$year)] <- mode_impute(retailc$year)
```

```{r}
retailc$ratings[is.na(retailc$ratings)]<- mean (retailc$ratings, na.rm=TRUE)
```

```{r}
dim(retailc)
```
### distinct: This is to delete duplicate
```{r}
retailc <- distinct(retailc)
```

```{r}
retailc <- retailc %>% distinct()
```

### Dim is count row and coloumn 
```{r}
dim(retailc)
```
```{r}
sum(is.na(retailc))
```
### this is to change the headlines to small letters 
```{r}
retailc <- clean_names(retailc)
```

### this is check for duplicate
```{r}
sum(duplicated(retailc))
```

```{r}
summary(retailc)
```
### This is to check the first 10 rows
```{r}
retailc
```


### Analysis pharse/ Exporatory Data Analysis (EDA)
```{r}
retailc %>%
  group_by(customer_segment) %>%
  summarize(total_purchase = sum (total_purchases),
            total_amount = sum(total_amount))
```

### THIS is to check for null values in a column
```{r}
sum(is.na(retailc$customer_segment))
```
### this is to check the grouping or category for the coloumn 
```{r}
unique(retailc$customer_segment)
```

###  this is use to remove or ignore null values in our coloumn (sustomer segement), na.rm = TRUE means if null is true
```{r}
filterer <- retailc%>%
  filter(!is.na(customer_segment))

filterer %>%
  group_by(customer_segment) %>%
  summarize(total_purchase = sum (total_purchases, na.rm = TRUE),
            total_amount = sum(total_amount, na.rm = TRUE))
```
```{r}
retailc %>%
  group_by(product_category) %>%
  summarize(total_purchase = sum (total_purchases),
            total_amount = sum(total_amount))
```

### 2. which product and the most popular among different customer segment?
```{r}
retailc %>%
  group_by(customer_segment) %>%
 count(product_category)
```

```{r}
retailc %>%
  group_by(product_category) %>%
  summarize(total_purchase = sum (total_purchases),
            total_amount = sum(total_amount)) %>%
  
  ### (this is arrange the column, to know what is the highest and lowest)
arrange (desc (total_purchase))
```

```{r}
#Replace empty strings with NA ( This is use when you are having Null values in any coloumn )
retailc$product_category[retailc$product_category == ""] <- NA 
```

```{r}
retailc %>%
  group_by(month) %>%
  summarize(total_purchase = sum (total_purchases))
```

```{r}
# Ensure 'month' is a factor with levels ordered from january to December 
# mutate is to arrange stuffs or add a new 

retailc <- retailc%>%
  mutate( month = factor(month, levels = c("January", "February", "March",
                                           "April", "May", "June", "July",
                                           "August", "September", "October",
                                           "November", "December")))

#Group by ordered factor and summarize the data 
monthly_Summary <- retailc %>%
  group_by(month) %>%
  summarize(total_purchase = sum (total_purchases, na.rm = TRUE))

#Display the result
print(monthly_Summary)
```

```{r}
# Ensure 'month' is a factor with levels ordered from january to December 
# mutate is to arrange stuffs or add a new 

retailc <- retailc%>%
  mutate( month = factor(month, levels = c("January", "February", "March",
                                           "April", "May", "June", "July",
                                           "August", "September", "October",
                                           "November", "December" )))

#Group by ordered factor and summarize the data 
monthly_Summary <- retailc %>%
  group_by(month) %>%
  summarize(total_purchase = sum (total_purchases, na.rm = TRUE))

#Display the result
print(monthly_Summary)
```

```{r}
packageVersion("ggplot2")
```

```{r}
summary(retailc)
```


### Visualization 
# Customer Segment Distribution PLOT 
# ggplot is used for visualization 
### geom_bar means bar chart while geom_line means line chart 
### theme means the colour of the backgroung 
### labs is the chart title with X & Y titles as well 
### aes specifies the coloumnin the df we're working with 
```{r}
ggplot(retailc, aes(x = Customer_Segment)) + 
  geom_bar() + 
  labs(title = "Customer Segments Distribution", X = "Segment", y = " Count")
```
```{r}
summary(retailc)
```
### Plot for Total Purchases by Customer Segment 

```{r}
ggplot(retailc, aes(x = customer_segment, y = total_purchases, fill = customer_segment)) +
  geom_bar(stat = "identity") +
  labs(title ="Total Purchases by Customer Segment", x = "customer_segment", y = "Total Purchases") +
  theme_minimal()
```
# Plot for Total Sales by Product Category
```{r}
ggplot(retailc, aes(x = product_category, y = total_purchases, fill = product_category)) +
geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Sales by Product Category", x = "Product Brand", y = "Total Sales") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Sales Trends Over Time 
```{r}
ggplot(retailc, aes(x = year, y = total_amount)) + 
  geom_line() + 
  labs(title = "Sales Trends Over Time", X = "Date", y = " Total Sales")
```

  